### 2016-04-22 manuscript-body.tex
Continue comparing Pea & Kurland with Bonar & Soloway

What's curious about Pea's (1986) comments on Bonar \& Soloway (1983) is
that Pea drew different conclusions from the data than Bonar and Soloway
did. Pea (1986) emphasized that when the student thought two
semantically-equivalent assignment statements in PASCAL were different,
it was a problem of egocentrism: ``students assume that there is
\emph{more} of their meaning for what they want to accomplish in the
program than is actually present in the code they have written'' (p.
30). In other words, Pea treated the data as a fairly clear example of a
class of bugs. Specifically, he saw a bug class in which students simply
assumed the interpreter or runtime could infer the code author's shades
of intent.

Bonar and Soloway (1983), by contrast, were less quick to make
inferences either about the nature of the bug or about the intervention
it entailed. Rather than jump to definitive conclusions, they were
circumspect:

It is not clear exactly how to react to the bugs we have uncovered in
novice understanding of programming. In some cases it may be appropriate
to design new languages or constructs. Often, better instruction would
take care of the problem. The intent of our studies is to better
understand the source of the mismatches and misconceptions that cause
novice bugs. Only once a bug is uncovered and understood are we ready to
create a remedy for that bug. (Bonar \& Soloway, 1983, p. 12)

That the authors would even consider developing new constructs or
languages is a noteworthy distinction. Rather than assume in toto that
students with ``bugs'' had wrong knowledge, the authors instead suppose
cognitive bugs have plausible origins worth designing around. Moreover,
they treat students' divergence from canon as \emph{an opportunity for
research to learn from students}. Precisely because students saw
functional differences (cf., diSessa, 1986) in semantically-equivalent
PASCAL statements, Bonar and Soloway (1983) reflect that perhaps
programming languages should be more expressive:

We find it quite interesting that novices seem to understand the role or
strategy of statements more clearly than the standard semantics. Such
roles discussed here include ``counter variable,'' ``running total
variable,'' ``running total loop,'' and ``first, then rest loop''. (See
Soloway et al {[}1982b{]} for a detailed discussion of novice looping
strategies.) Much work in programming languages is concerned with
allowing a programmer to more accurately express his or her intentions
in the program. Perhaps we can learn something from novices here - our
programming systems should support recording the roles the programmer
intends for various statements and variables. (Bonar \& Soloway, 1983,
p. 12)

Again, what's noteworthy here is that rather than treating students'
non-canonical views as a burden for instruction, Bonar and Soloway
instead see them as an opportunity for programming language designers to
make languages better.

@@

The experience and understanding of a novice are available for analysis.
In particular, our results suggest that the knowledge people bring from
natural language has a key effect on their early programming efforts.
Our work suggests that we need serious study of the knowledge novices
bring to a computing system. For most computerized tasks there is some
model that a novice will use in his or her first attempts. We need to
understand when it is appropriate to appeal to this model, and, when
necessary, how to move a novice to some more appropriate model. (Bonar
\& Soloway, 1983, p. 13)

I assume Bonar and Soloway structured their final line deliberately. If
so, their phrasing has three consequences:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Appealing to novice's existing models gets precedence. That is,
  understanding how to leverage novices' existing knowledge comes first
  in research.
\item
  Changing the models novices have comes next, and ``when necessary.''
\item
  They speak of ``how to move a novice to some more appropriate model,''
  which does not necessarily entail the removal or destruction of
  novice's existing models.
\end{enumerate}

>>>>

What's curious about Pea and Kurland's \cite{pea_languageindependent_1986} comments on Bonar and Solowa \cite{bonar_uncovering_1983} is
that \emph{Pea \cite{pea_languageindependent_1986} drew different conclusions from the data than did Bonar and Soloway
}. Pea emphasized that when the student thought two
semantically-equivalent assignment statements in PASCAL were different,
it was a problem of egocentrism: ``students assume that there is
\emph{more} of their meaning for what they want to accomplish in the
program than is actually present in the code they have written'' \cite{pea_languageindependent_1986}. In other words, Pea treated the data as a fairly clear example of a
class of cognitive bugs. Specifically, he saw a bug class in which students simply
assumed the interpreter or runtime could infer programmer's shades
of intent.

Bonar and Soloway (1983), by contrast, were less quick to make
inferences either about the nature of the bug or the intervention
it entailed. Rather than jump to definitive conclusions, they were
circumspect:

\begin{quote}
  It is not clear exactly how to react to the bugs we have uncovered in
  novice understanding of programming. In some cases it may be appropriate
  to design new languages or constructs. Often, better instruction would
  take care of the problem. The intent of our studies is to better
  understand the source of the mismatches and misconceptions that cause
  novice bugs. Only once a bug is uncovered and understood are we ready to
  create a remedy for that bug. \cite{bonar_uncovering_1983}
\end{quote}

That Bonar and Soloway would even consider developing new constructs or
languages is a noteworthy distinction. Rather than assume wholesale that
students with ``bugs'' had wrong knowledge, the authors instead suppose
cognitive bugs have plausible origins worth designing around. Moreover,
they treat students' divergence from canon as \emph{an opportunity for
research to learn from students}. Precisely because students saw
functional differences (cf., diSessa, 1986) in semantically-equivalent
PASCAL statements, Bonar and Soloway (1983) reflect that perhaps
programming languages should be more expressive:

\begin{quote}
  We find it quite interesting that novices seem to understand the role or
  strategy of statements more clearly than the standard semantics. Such
  roles discussed here include ``counter variable,'' ``running total
  variable,'' ``running total loop,'' and ``first, then rest loop''. (See
  Soloway et al {[}1982b{]} for a detailed discussion of novice looping
  strategies.) Much work in programming languages is concerned with
  allowing a programmer to more accurately express his or her intentions
  in the program. Perhaps we can learn something from novices here - our
  programming systems should support recording the roles the programmer
  intends for various statements and variables. \cite{bonar_uncovering_1983}
\end{quote}

Again, what's noteworthy is that rather than treating students'
non-canonical views as a burden for instruction, Bonar and Soloway
instead see them as an opportunity for programming language designers to
make languages better.

@@

\begin{quote}
  The experience and understanding of a novice are available for analysis.
  In particular, our results suggest that the knowledge people bring from
  natural language has a key effect on their early programming efforts.
  Our work suggests that we need serious study of the knowledge novices
  bring to a computing system. For most computerized tasks there is some
  model that a novice will use in his or her first attempts. We need to
  understand when it is appropriate to appeal to this model, and, when
  necessary, how to move a novice to some more appropriate model. (Bonar
  \& Soloway, 1983, p. 13)
\end{quote}

We assume Bonar and Soloway structured the final line of that quote deliberately. If
so, their phrasing has three consequences:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Appealing to novice's existing models gets precedence. That is,
  understanding how to leverage novices' existing knowledge comes first
  in research.
\item
  Changing the models novices have comes next, and ``when necessary.''
\item
  They speak of ``how to move a novice to some more appropriate model,''
  which does not necessarily entail the removal or destruction of
  novice's existing models.
\end{enumerate}
